TBD

### 2023-03-09 outputs/paper/paper3.qmd
Completed first pass at financial wellness x demographic factors; reorganized paper structure

## Data

@@

### Measurement Bias

@@

@@

Higher degree - more financially satisfied

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(marital), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = marital)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~marital)
```

Those born in US are less satisfied?

# # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
# survey_data_2021 |>
#   filter(!is.na(born), !is.na(financial_satisfaction)) |>
#   ggplot(aes(x = financial_satisfaction, group = born)) +
#   geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
#   scale_y_continuous(labels=scales::percent) +
#   ylab("Percentage of response") +
#   xlab("Financial Satisfaction") +
#   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
#   theme(legend.position="none") +
#   facet_grid(~born)
```

Older cohorts are more satisfied with their financial situation

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = age_band)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~age_band)

@@

Some differences based on working status

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(work_status), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = work_status)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~work_status)
```

Sexual orientation - not too much of a pattern?

# # Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(marital), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = marital)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~marital)
```

Social status - interestingly, upper class is less satisfied than upper middle class?

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(social_class1), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = social_class1)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~social_class1)
```

Financial satisfaction - does money make you more financially satisfied? oh yea.

```{r}
#| message: false
#| echo: false
#| warning: false
#| 
survey_data_2021 |>
  filter(!is.na(family_income_bands), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = family_income_bands)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~family_income_bands)
```

@@

@@

@@

>>>>

## General Social Survey

@@

@@

survey_data_2021 |>
  filter(!is.na(degree), !is.na(financial_satisfaction)) |>
  group_by(degree, financial_satisfaction) |>
  count() |>
  ungroup() |>
  group_by(degree) |>
  mutate(pct = percent(n / sum(n), digit = 1)) |>
  ggplot(aes(x = degree, y = pct, fill = financial_satisfaction)) +
  geom_bar(stat="identity", position="dodge") +
  scale_y_continuous(labels=scales::percent) +
  theme_minimal() +
  theme(legend.position="bottom") + 
  labs(
    x = "Financial Satisfaction",
    y = "Percentage of Responses",
    fill = "Age Group"
  )

survey_data_2021 |>
  filter(!is.na(degree), !is.na(financial_change)) |>
  group_by(degree, financial_change) |>
  count() |>
  ungroup() |>
  group_by(degree) |>
  mutate(pct = percent(n / sum(n), digit = 1)) |>
  ggplot(aes(x = degree, y = pct, fill = financial_change)) +
  geom_bar(stat="identity", position="dodge") +
  scale_y_continuous(labels=scales::percent) +
  theme_minimal() +
  theme(legend.position="bottom") + 
  labs(
    x = "Change in Financial Situation",
    y = "Percentage of Responses",
    fill = "Age Group"
  )
```

The 2021 GSS data shows that having a higher level of education is correlated with higher financial satisfaction. in @fig-2021-degree-financial-1 there is a trend of decreasing percentages of answers for "not satisfied at all" as level of education increases, as well as an increasing percentages of responses for "pretty well satisfied" as the level of education increases. This finding is supported by the survey results for respondents' perceptions on the changes in the financial situation (@fig-2021-degree-financial-1). As the level of education increases, there is a noticeable increase in the percentages of answers for "getting better" in their financial situation. Stay in school kids, it will help you financially in the long run.

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-2021-income-financial
#| fig-cap: "Intersection of financial attitudes with family income (2016)"
#| fig-subcap: ["Financial Satisfaction", "Change in Financial Situation"]
#| layout-ncol: 2

survey_data_2021 |>
  filter(!is.na(financial_satisfaction), !is.na(income)) |>
  group_by(financial_satisfaction, income) |>
  count() |>
  ungroup() |>
  group_by(income) |>
  mutate(Percentage = n / sum(n)) |>
  ggplot(aes(x = financial_satisfaction, y = income, fill = Percentage)) +
  geom_tile() + 
  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs (
    x = "Financial Satisfaction",
    y = "Family Income in 2016",
  )

survey_data_2021 |>
  filter(!is.na(financial_change), !is.na(income)) |>
  group_by(financial_change, income) |>
  count() |>
  ungroup() |>
  group_by(income) |>
  mutate(Percentage = n / sum(n)) |>
  ggplot(aes(x = financial_change, y = income, fill = Percentage)) +
  geom_tile() + 
  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs (
    x = "Comparing Financial Situation to Others",
    y = "Self-Identified Social Class",
  )

@@

As a hypothesis, we would expect attitudes towards financial wellness is highly correlated with respondents' financial income. This is supported in the our data, as @fig-2021-income-financial-1 shows that there is a high percentage of respondents with family income less than \$30K reporting not being satisfied with their current financial situation. On the flip side, the majority of respondents making \$170K and above are pretty well satisfied with their current situation. There is also the general trend that as the family income increases, the percentage of being satisfied with their current situation also increases. Looking at the changes in financial situation, the data shows that those making less than \$30K are getting worse or staying the same, while the majority of respondents making more than \$110K are getting better in their financial situations (@fig-2021-income-financial-2).

@@

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(social_class1), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = social_class1)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~social_class1)

survey_data_2021 |>
  filter(!is.na(social_class), !is.na(income)) |>
  group_by(social_class, income) |>
  count() |>
  ungroup() |>
  group_by(income) |>
  mutate(Percentage = n / sum(n)) |>
  ggplot(aes(x = social_class, y = income, fill = Percentage)) +
  geom_tile() + 
  scale_fill_gradient2(low="white", high="darkgreen", guide="colorbar") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs (
    x = "Comparing Financial Situation to Others",
    y = "Self-Identified Social Class",
  )
```

@@

## Marginalization against segments of the population

Rich and richer; poor getting poorer

## Newer generations not doing as well financially

Not trending well over time

##

## Fleishman Effect

Upper class in their own bubble

@@

Future studies - incorporate more survey responses

More correlations between factors
